Clause Size Reduction with \alluip Learning

### 2020-05-08 paper.tex
improved alg description

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\oneuip}{1-UIP\xspace}
\newcommand{\alluip}{all-UIP\xspace}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\stablealluip}{\textit{stable-alluip}}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
\newcommand{\allUipAct}{\textit{alluip-active}}
\newcommand{\allUipIn}{\textit{alluip-inclusive}}
\newcommand{\allUipEx}{\textit{alluip-exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\type}{\textit{type}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

@@

The \alluip scheme will tend to produce small clauses since the
clauses contain at most one literal per decision level. However, the
\alluip clause can have much higher \LBD. Since $\LBD$ is more
important than size, our approach is to use \alluip learning when, and
only when, it succeeds in reducing the size of the clause
\emph{without increasing its \LBD}. The \alluip scheme first computes
the \oneuip clause when it reduces the deepest level to a single UIP
literal. It then proceeds to reduce the shallower levels (see
\alluip's for loop in Figure~\ref{fig:cl_schemes}). So our approach
will start with the \oneuip clause and then try to apply \alluip
learning to reduce other levels to single literals. As noted above,
clause minimization is orthogonal to our approach, so we also first
apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
to the \oneuip clause. That is, our algorithm will start with the
clause that most \sat solvers learn from a conflict, a minimized
\oneuip clause. Our approach is specified in Algorithm~\ref{alg:i-uip}
($\stablealluip$).

\begin{algorithm}[!t]
\caption{\stablealluip}\label{alg:i-uip}
\begin{algorithmic}[1]
    \Require $C_1$ is minimized \oneuip clause
    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are globals used to
    configure $\tgap$,   \type{} a global determining how the
    introduction of new decision levels is handled.

    \Procedure{\stablealluip}{$C_1, \assertionTrail$, \type} \State
    \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return}
    $C_1$\label{ln:control} \State $\ntries$++\label{ln:try} \State
    $C_i \gets C_1$ \State $\dlevels \gets\mbox{}$ decision levels in
    $C_i$ in descending order \Comment{These never change}
    \For{($i=1$; $i < |\dlevels|$; $i$++)} \Comment{skip the deepest
      level $\dlevels[0]$} \State $C_i\gets\mbox{}$\tryuiplevel($C_i$,
    $\dlevels[i]$, \type) \Comment{Try to reduce this level to UIP}
    \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq
      \dlevels[i]\}\bigr| + (|\dlevels| - (i+1)) \geq
      |C_1|$}\label{ln:earlyTerminate} \State \textbf{return}
    $C_1$. \Comment{can't generate smaller clause} \EndIf \EndFor
    \State \textbf{if} ($\type = \allUipPure$)
    $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
    %\State $\ntries$++\label{ln:try}% 
    \State \textbf{if} ($|C_i| < |C_1|$) $\nsuc$++, \textbf{return} $C_i$ 
           \label{ln:returnmin}\label{ln:suc}
    \State \textcolor{gray}{$\triangleright\mbox{}$ Use next line instead for} \allUipAct
    \State \textcolor{gray}{
           \textbf{if} ($|C_i| < |C_1| 
                        \land (\code{AvgVarAct}(C_i) > \code{AvgVarAct}(C_1))$)
                       $\nsuc$++, \textbf{return} $C_i$} \label{ln:activity}
    \State \textbf{return} $C_1$
%    \State \textbf{return} \textbf{if} $|C_i| < |C_1|$) 
%           \textbf{then}  $C_i$ \textbf{else}  $C_1$
\EndProcedure
\Statex 
\Procedure{\tryuiplevel}{$C_i$, $i$, \type}\Comment{
  \parbox[t]{3in}{
    Do not add new decision levels, \type{} determines
    what to do if new decision levels would be introduced
  }
}
    \State $\ctry = C_i$
    \State $L_i = \left\{\ell \left| \ell \in \ctry 
                         \land \dlevel{l} = i\right.\right\}$\label{ln:litsati1}
    \While{$|L_i| > 1$}
        \State $p \gets\mbox{}$ \textbf{remove} lit with the highest trail index from $L_i$ 
        \If{($\exists q \in \reason{\lnot p}.\, \dlevel{q} \not \in \mathit{decLvls}$)}\label{ln:newdlevel}
        \Comment{Would add new decision levels}
             \If{($\type = \allUipPure$)}
                 \State \textbf{return} $C_i$ \Comment{Abort, can't UIP this level}
             \ElsIf{($\type = \allUipMin$)}
                 \State \textbf{continue} \Comment{Don't try to resolve away $p$}
             \EndIf
        \Else\label{ln:nonewdlevel}
            \State $\ctry \gets \ctry \resolve \reason{\lnot p}$
            \State $L_i = L_i \cup \{\ell\,|\, 
                                     \ell\in\reason{\lnot p} \land \ell\neq \lnot p \land \dlevel{\ell} = i
                                    \}$\label{ln:litsati2}
        \EndIf
     \EndWhile
    \State \textbf{return} $\ctry$
\EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:i-uip} attempts to reduce each level in the \oneuip
clause to a single literal (a UIP). The \oneuip clause has only one
literal at its deepest level, so that level, $\dlevels[0]$, can be
skipped. The subroutine \tryuiplevel is used to reduce each level. It
uses trail resolutions to achieve this, subject to the constraint that
no new decision levels can be introduced in the clause. In particular,
\tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
(line~\ref{ln:litsati1}), in order of decreasing trail index, until
only one literal at level $i$ remains. If the resolution step will not
introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
performed updating $\ctry$. In addition, all new literals at level $i$
are added to $L_i$. These are the literals at level $i$ in the reason
clause.

On the other hand, if the resolution step would introduce new decision
levels (line~\ref{ln:newdlevel}) then there are two options.  The
first option we call $\allUipPure$. With $\allUipPure$ we abort our
attempt to UIP this level and return the clause with level $i$
unchanged. In the second option, called $\allUipMin$, we continue
without performing the resolution, \textit{keeping} the current
literal $p$ in $\ctry$. $\allUipMin$ then continues to try to resolve
away the other literals in $L_i$ (note that $p$ is no longer in $L_i$)
until $L_i$ is reduced to a single literal. Hence, $\allUipMin$ can
leave multiple literals at level $i$---all of those with reasons
containing new levels along with one other.\footnote{Since the sole
  remaining literal $u\in L_i$ is at a lower trail index than all of
  the other literals there is no point in trying to resolve away
  $u$---either it will be the decision literal for level $i$ having no
  reason, or its reason will contain at least one other literal at
  level $i$.} Observe that the number of literals at level $i$ can not
be increased after processing it with \allUipPure. \allUipMin can,
however, potentially increase the number of literals at level $i$. In
resolving away a literal $l$ at level $i$, more literals might be
introduced into level $i$, and some of these might not be removable by
\allUipMin if their reasons contain new levels. However, both
\allUipPure and \allUipMin can increase the number of literals at
levels less that $i$ as new literals can be introduced into those
levels when the literals at level $i$ are resolved away. These added
literals at the lower levels might not be removable from the clause,
and thus both methods might yield a longer clause than the inputted
\oneuip clause.

After trying to UIP each level the clause $C_i$ is obtained. If we
were using $\allUipPure$ we can once again apply minimization
(line~\ref{ln:puremin}). Recursive clause minimization
\cite{DBLP:conf/sat/SorenssonB09} would be useless for the $\allUipMin$
clause as all but one literal of each level introduces a new level and
thus cannot be recursively removed.\footnote{Other more powerful
  minimization techniques could still be applied.}

Finally, the algorithm returns the shorter of the initial \oneuip clause
and the newly computed clause (line~\ref{ln:returnmin}). Because we
return the newly computed clause only when it is shorter, an early
termination test can be used (line~\ref{ln:earlyTerminate}). After the
algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
the literals at those levels will not be further changed. Furthermore,
we know that the best that can be done is to reduce the remaining
$|\dlevels| - (i+1)$ levels down to a single literal each. These two
observations give a lower bound on the size of the resulting clause,
and if that lower bound is as large as the size of the initial \oneuip
clause we can terminate and return the initial \oneuip clause.  This
termination condition does not, however, account for the final clause
minimization (line~\ref{ln:puremin}) of \allUipPure, and it could
prematurely terminate processing that would otherwise yield a shorter
clause. However, we discovered such false negative is rare, and
applying early termination on \allUipPure generally improves
performance.
%\allUipPure cannot trigger this early termination
%condition (it cannot increase the size of any processed level), but
%\allUipMin can.

\paragraph{$\tgap$:}
$\stablealluip$ can produce significantly smaller clause. However, when it
does not yield a smaller clause, the cost of the additional resolution
steps can hurt the solver's performance. Since resolution cannot
reduce a clause's $\LBD$, the maximum size reduction obtainable from
$\stablealluip$ is the difference between the \oneuip clause's size and its
$\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$ is small,
applying $\stablealluip$ is unlikely to be cost effective. Our approach is
to dynamically set a threshold on $\gap(C_1)$, $\tgap$, such that when
$\gap(C_1) < \tgap$ we do not attempt to reduce the clause
(line~\ref{ln:control}). Initially, $\tgap = 0$, and we count the
number of times $\stablealluip$ is attempted ($\ntries$) and the number of
times it successfully yields a shorter clause ($\nsuc$)
(line~\ref{ln:try} and \ref{ln:suc}). On every restart if the success
rate since the last restart is greater than 80\%, we decrease $\tgap$ by one (not allowing it to
become negative), and if it is less than 80\% we increase $\tgap$
by one.

\begin{example}
    Consider the trail $\assertionTrail = \ldots$, $\ell_1$, $a_2$, $b_2$,
    $c_2$,$d_2$, $\ldots$, $\ldots$, $e_5$, $f_5$, $g_5$,
    $h_6$, $i_6$, $j_6$, $k_6$, $\ldots$, $m_{10}$,
    $\ldots$ where the subscript indicates the decision level of each
    literal and the literals are in order of increasing trail
    index. Let the clauses
    $C_x$ denote the reason clause for literal $x_i$ with 
    \[\begin{array}{|l|l|l|}\hline
      C_{a} = \varnothing 
      & C_{b} =  (b_2, \lnot \ell_3, \lnot a_2)
      & C_{c} =  (c_2, \lnot a_2, \lnot b_2) \\\hline
      C_{d} =  (d_2, \lnot b_2, \lnot c_2) 
      & C_{\ell} =\varnothing
      & C_{e} =\varnothing \\\hline
      C_{f} =  (f_5, \lnot e_5,\lnot \ell_1)
      & C_{g} = (g_5, \lnot a_2, \lnot f_5)
      & C_{h} =\varnothing \\\hline
      C_{i} =  (i_6, \lnot e_5, \lnot h_6)
      & C_{j} = (j_6, \lnot f_5, \lnot i_6 )
      & C_{k} = (k_6, \lnot f_5, \lnot j_6) \\\hline
      \end{array}
  \]
  Suppose \oneuip learning yields the clause
  $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{m_{10}}$
  is the UIP from the conflicting level. $\stablealluip$ first tries to find
  the UIP for level 6 by resolving $C_1$ with $C_k$, $C_j$
  and then $C_i$ producing the clause
  $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$,
  $\neg{f_5}$, $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where
  $\neg{h_6}$ is the UIP for level 6.

@@

\subsubsection{$\allUipAct$: Clauses with Active Variables.}
\label{sec:active} $\stablealluip$ learning might
introduce literals with low variable activity into $C_i$. Low activity
variables are variables that have had low recent participation in
clause learning. Hence, clauses with variables of low activity might
not be as currently useful to the solver. Our variant $\allUipAct$,
shown as the optional line~\ref{ln:activity} in
Algorithm~\ref{alg:i-uip}, computes the average variable activity of
the newly produced \alluip clause $C_i$ and the original \oneuip
clause $C_1$. The new clause $C_i$ will be returned only if it is both
smaller and has higher average variable activity than the original
\oneuip clause. There are, of course, generalizations of this approach
where one has a weighted trade-off between these factors that allows
preferring the new clause when it has large gains in one metric even
though it has small losses in the other. We did not, however,
experiment with such generalizations.

>>>>

\newcommand{\fs}[1]{\fontsize{#1}{#1}\selectfont}
\newcommand{\fss}[2]{\fontsize{#1}{#2}\selectfont}
\newcommand{\sat}{SAT\xspace}
\newcommand{\code}[1]{\text{#1}}
\newcommand{\assertionTrail}{\trail}
\newcommand{\oneuip}{1-UIP\xspace}
\newcommand{\alluip}{all-UIP\xspace}
\newcommand{\LBD}{\text{LBD}\xspace}
\newcommand{\gap}{\text{gap}}
\newcommand{\tgap}{t_{\mathit{gap}}}
\newcommand{\ntries}{n_{\mathit{tries}}}
\newcommand{\nsuc}{n_{\mathit{succ}}}
\newcommand{\stablealluip}{\textit{stable-alluip}\xspace}
\newcommand{\tryuiplevel}{\textit{try-uip-level}\xspace}
\newcommand{\allUipPure}{\textit{pure-alluip}\xspace}
\newcommand{\allUipMin}{\textit{min-alluip}\xspace}
\newcommand{\allUipAct}{\textit{alluip-active}}
\newcommand{\allUipIn}{\textit{alluip-inclusive}}
\newcommand{\allUipEx}{\textit{alluip-exclusive}}
\newcommand{\MapleBase}{\textit{MapleCOMSPS\_LRB}}
\newcommand{\MapleSeven}{\textit{MapleLCMDist}}
\newcommand{\MapleNine}{\textit{MapleLCMDiscChronoBT-DL-v3} }
\newcommand{\MapleEight}{\textit{MapleLCMDiscChronoBT}}
\newcommand{\MapleNineShort}{\textit{MapleCB-DL} }
\newcommand{\expSAT}{\textit{expMaple-CM-GCBumpOnlyLRB} }
\newcommand{\expSATShort}{\textit{expMaple} }
\newcommand{\MapleIUIPPure}{\text{Maple-\allUipPure}}
\newcommand{\MapleIUIMin}{\text{Maple-\allUipMin}}
\newcommand{\MapleEightShort}{\textit{MapleCB}}
\newcommand{\Set}[2]{\{\,#1\mid#2\,\}}
\newcommand{\Array}[2]{[\,#1\mid#2\,]}
\newcommand{\trail}{\ensuremath{\mathcal{T}}}
\newcommand{\trailIdx}[1]{\ensuremath{\iota(#1)}}
\newcommand{\dlevel}[1]{\ensuremath{\mathit{decLvl}(#1)}}
\newcommand{\dlevels}{\ensuremath{\mathit{decLvls}}}
\newcommand{\var}{\text{var}}
\newcommand{\true}{\textsc{true}\xspace}
\newcommand{\false}{\textsc{false}\xspace}
\newcommand{\reason}[1]{\ensuremath{\mathit{reason}(#1)}}
\newcommand{\resolve}{\bowtie}
\renewcommand{\implies}{\rightarrow}
\newcommand{\ctry}{C_{\mathit{try}}}
\newcommand{\config}{\textit{config}}
\newcommand{\unmarked}{\textbf{unmarked}}
\newcommand{\mrk}{\textbf{mark}}
\setlength{\theorempreskipamount}{3pt}
\setlength{\theorempostskipamount}{3pt}

@@

The \alluip scheme will tend to produce small clauses since the
clauses contain at most one literal per decision level. However, the
\alluip clause can have much higher \LBD. Since $\LBD$ is more
important than size, our approach is to use \alluip learning when, and
only when, it succeeds in reducing the size of the clause
\emph{without increasing its \LBD}. The \alluip scheme first computes
the \oneuip clause when it reduces the deepest level to a single UIP
literal. It then proceeds to reduce the shallower levels (see
\alluip's for loop in Figure~\ref{fig:cl_schemes}). So our approach
will start with the \oneuip clause and then try to apply \alluip
learning to reduce other levels to single literals. As noted above,
clause minimization is orthogonal to our approach, so we also first
apply standard clause minimization \cite{DBLP:conf/sat/SorenssonB09}
to the \oneuip clause. That is, our algorithm $\stablealluip$,
Algorithm~\ref{alg:i-uip}, starts with the clause that most \sat
solvers learn from a conflict, a minimized \oneuip clause.

\begin{algorithm}[!t]
\caption{\stablealluip}\label{alg:i-uip}
\begin{algorithmic}[1]
    \Require $C_1$ is minimized \oneuip clause
    \Require \config{} a set of configuration parameters to give different versions \stablealluip.
    \Require $\tgap\geq 0$ is a global parameter, $\ntries$ and $\nsuc$ are used to
         dynamically adjust $\tgap$

    \Procedure{\stablealluip}{$C_1, \assertionTrail$}
    \State \textbf{if} ($|C_1| - \LBD(C_1) < \tgap$) \textbf{return} $C_1$\label{ln:control} 
    \State $\ntries$++\label{ln:try} 
    \State $C_i \gets C_1$
    \State $\dlevels \gets\mbox{}$ decision levels in $C_1$ in descending order\label{ln:init_loop}
        \Comment{These never change}
    \For{($i=1$; $i < |\dlevels|$; $i$++)}\label{ln:begin_level_loop}
         \Comment{skip the deepest level $\dlevels[0]$}
        \State $C_i\gets\mbox{}$\tryuiplevel($C_i$, $\dlevels[i]$) 
            \Comment{Try to reduce this level to UIP}
        \If{$\bigl|\{\ell\,|\,\ell \in C_i \land \dlevel{\ell} \geq
            \dlevels[i]\}\bigr| + (|\dlevels| - (i+1)) \geq
            |C_1|$}\label{ln:earlyTerminate}
            \State \textbf{return} $C_1$ 
                \Comment{can't generate smaller clause}\label{ln:end_level_loop}
        \EndIf
   \EndFor
   \If{$\mathbf{\allUipPure} \in \config$}
       \State $C_i \gets \mathbf{minimize}(C_i)$\label{ln:puremin}
   \EndIf
   \If{$\bigl(|C_i| < |C_1| \land 
       \mathbf{\allUipAct}\in\config\implies(\code{AvgVarAct}(C_i) > \code{AvgVarAct}(C_1))$
      \label{ln:activity}}
        \State $\nsuc$++, \textbf{return} $C_i$  \label{ln:returnmin}\label{ln:suc}
        \Comment{$C_i$ is smaller than the input clause}
   \Else \State \textbf{return} $C_1$
   \EndIf
%    \State \textbf{return} \textbf{if} $|C_i| < |C_1|$) 
%           \textbf{then}  $C_i$ \textbf{else}  $C_1$
\EndProcedure
\Statex 
\Procedure{\tryuiplevel}{$C_i$, $i$}
    \Comment{Do not add new decision levels}
    \State $\ctry = C_i$
    \State $L_i = \left\{\ell \left| \ell \in \ctry 
                         \land \dlevel{l} = i\right.\right\}$\label{ln:litsati1}
    \While{$|L_i| > 1$}
        \State $p \gets\mbox{}$ \textbf{remove} lit with the highest trail index from $L_i$ 
        \If{($\exists q \in \reason{\lnot p}.\, \dlevel{q} \not \in \mathit{decLvls}$)}\label{ln:newdlevel}
        \Comment{Would add new decision levels}
             \If{($\mathbf{\allUipPure}\in \config$)}
                 \State \textbf{return} $C_i$ \Comment{Abort, can't UIP this level}
             \ElsIf{($\mathbf{\allUipMin}\in\config$)}
                 \State \textbf{continue} \Comment{Don't try to resolve away $p$}
             \EndIf
        \Else\label{ln:nonewdlevel}
            \State $\ctry \gets \ctry \resolve \reason{\lnot p}$
            \State $L_i = L_i \cup \{\ell\,|\, 
                                     \ell\in\reason{\lnot p} \land \ell\neq \lnot p \land \dlevel{\ell} = i
                                    \}$\label{ln:litsati2}
        \EndIf
     \EndWhile
    \State \textbf{return} $\ctry$
\EndProcedure
\end{algorithmic}
\end{algorithm}

Algorithm~\ref{alg:i-uip} tries to compute a clause shorter than the
inputted \oneuip clause $C_1$. If a clause shorter than $C_1$ cannot
be computed the routine returns $C_1$ unchanged. Line~\ref{ln:control}
uses the parameter $\tgap$ to predict if Alg.~\ref{alg:i-uip} will be
successful in producing a shorter clause. This predication is
described below. If the prediction is negative $C_1$ is immediately
returned and Alg.~\ref{alg:i-uip} is not attempted. Otherwise, a copy
of $C_1$ is made in $C_i$ and $\ntries$, which counts the number of
times Alg.~\ref{alg:i-uip} is attempted, is incremented.

Then the decision levels of $C_1$ are computed and stored in
$\dlevels$ in order from largest to lowest. The for loop of
lines~\ref{ln:begin_level_loop}--\ref{ln:end_level_loop} is then
executed for each decision level $\dlevels[i]$. In the loop the
subroutine \tryuiplevel tries to reduce the set of literals at
$\dlevels[i]$ down to a single UIP literal using a sequence of trail
resolutions. Since $C_1$ is a \oneuip clause $\dlevels[0]$ (the
deepest level) already contains only one literal, so we can start at
$i=1$.

After the call to \tryuiplevel a check (line~\ref{ln:earlyTerminate})
is made to see if we can abort further processing. At this point the
algorithm has finished processing levels $\dlevels[0]$--$\dlevels[i]$
so the literals at those levels will change. Furthermore, we know that
the best that can be done from this point on is to reduce the
remaining $|\dlevels| - (i+1)$ levels down to a single literal
each. Hence, adding these two numbers gives a lower bound on the size
of the final computed clause. If that lower bound is as large as the
size of the initial \oneuip clause we can terminate and return the initial
\oneuip clause.

After processing all decision levels, if \tryuiplevel is using the
\allUipPure configuration, additional reduction in the clause size
might be achieved by an another round of clause minimization
(line~\ref{ln:puremin}). Finally, if the newly computed clause $C_i$
is smaller that the input clause $C_1$ it is returned. Otherwise the
original clause $C_1$ is returned. Additionally, if the configuration
$\allUipAct$, described in Section~\ref{sec:active}, is being used,
then we also require that the average activity level of the new clause
$C_i$ be larger than $C_1$ before we can return the new clause $C_i$.

\tryuiplevel($C_i$, $i$) attempts to resolve away the literals at
decision level $i$ in the clause $C_i$, i.e., those in the set $L_i$
(line~\ref{ln:litsati1}), in order of decreasing trail index, until
only one literal at level $i$ remains. If the resolution step will not
introduce any new decision levels (line~\ref{ln:nonewdlevel}), it is
performed updating $\ctry$. In addition, all new literals added to
$\ctry$ at level $i$ are added to $L_i$. These are the literals at
level $i$ in the reason clause.

On the other hand, if the resolution step would introduce new decision
levels (line~\ref{ln:newdlevel}) then there are two options.  The
first option we call $\allUipPure$. With $\allUipPure$ we abort our
attempt to UIP this level and return the clause with level $i$
unchanged. In the second option, called $\allUipMin$, we continue
without performing the resolution, keeping the current literal $p$ in
$\ctry$. $\allUipMin$ then continues to try to resolve away the other
literals in $L_i$ (note that $p$ is no longer in $L_i$) until $L_i$ is
reduced to a single literal. Hence, $\allUipMin$ can leave multiple
literals at level $i$---all of those with reasons containing new
levels along with one other.\footnote{Since the sole remaining literal
  $u\in L_i$ is at a lower trail index than all of the other literals
  there is no point in trying to resolve away $u$---either it will be
  the decision literal for level $i$ having no reason, or its reason
  will contain at least one other literal at level $i$.} Observe that
the number of literals at level $i$ can not be increased after
processing it with \allUipPure. \allUipMin can, however, potentially
increase the number of literals at level $i$. In resolving away a
literal $l$ at level $i$, more literals might be introduced into level
$i$, and some of these might not be removable by \allUipMin if their
reasons contain new levels. However, both \allUipPure and \allUipMin
can increase the number of literals at levels less that $i$ as new
literals can be introduced into those levels when the literals at
level $i$ are resolved away. These added literals at the lower levels
might not be removable from the clause, and thus both methods might
yield a longer clause than the inputted \oneuip clause.

After trying to UIP each level the clause $C_i$ is obtained. If we
were using $\allUipPure$ we can once again apply recursive clause
minimization (line~\ref{ln:puremin})
\cite{DBLP:conf/sat/SorenssonB09}, but this would be useless when
using $\allUipMin$ as all but one literal of each level introduces a
new level and thus cannot be recursively removed.\footnote{Other more
  powerful minimization techniques could still be applied.}

%\allUipPure cannot trigger this early termination
%condition (it cannot increase the size of any processed level), but
%\allUipMin can.

\noindent\textbf{$\tgap$}:\,\,
$\stablealluip$ can produce significantly smaller clauses. However,
when it does not yield a smaller clause, the cost of the additional
resolution steps can hurt the solver's performance. Since resolution
cannot reduce a clause's $\LBD$, the maximum size reduction obtainable
from $\stablealluip$ is the difference between the \oneuip clause's
size and its $\LBD$: $\gap(C_1) = |C_1| - \LBD(C_1)$. When $\gap(C_1)$
is small, applying $\stablealluip$ is unlikely to be cost
effective. Our approach is to dynamically set a threshold on
$\gap(C_1)$, $\tgap$, such that when $\gap(C_1) < \tgap$ we do not
attempt to reduce the clause (line~\ref{ln:control}). Initially,
$\tgap = 0$, and we count the number of times $\stablealluip$ is
attempted ($\ntries$) and the number of times it successfully yields a
shorter clause ($\nsuc$) (line~\ref{ln:try} and \ref{ln:suc}). On
every restart if the success rate since the last restart is greater
than 80\% (less than 80\%), we decrease (increase) $\tgap$ by one
not allowing it to become negative.

\vspace*{-3pt}
\begin{example}
    Consider the trail $\assertionTrail = \ldots$, $\ell_1$, $a_2$, $b_2$,
    $c_2$,$d_2$, $\ldots$, $\ldots$, $e_5$, $f_5$, $g_5$,
    $h_6$, $i_6$, $j_6$, $k_6$, $\ldots$, $m_{10}$,
    $\ldots$ where the subscript indicates the decision level of each
    literal and the literals are in order of increasing trail
    index. 
    \[\begin{array}{|l|l|l|}\hline
      C_{a} = \varnothing 
      & C_{b} =  (b_2, \lnot \ell_3, \lnot a_2)
      & C_{c} =  (c_2, \lnot a_2, \lnot b_2) \\\hline
      C_{d} =  (d_2, \lnot b_2, \lnot c_2) 
      & C_{\ell} =\varnothing
      & C_{e} =\varnothing \\\hline
      C_{f} =  (f_5, \lnot e_5,\lnot \ell_1)
      & C_{g} = (g_5, \lnot a_2, \lnot f_5)
      & C_{h} =\varnothing \\\hline
      C_{i} =  (i_6, \lnot e_5, \lnot h_6)
      & C_{j} = (j_6, \lnot f_5, \lnot i_6 )
      & C_{k} = (k_6, \lnot f_5, \lnot j_6) \\\hline
      \end{array}
  \]
  Let the clauses $C_x$, show above, denote the reason clause for
  literal $x_i$.  Suppose \oneuip learning yields the clause
  $C_1 = (\neg{m_{10}}$, $\neg{k_6}$, $\neg{j_6}$, $\neg{i_6}$,
  $\neg{h_6}$, $\neg{g_5}$, $\neg{d_2}$, $\neg{c_2})$ where
  $\neg{m_{10}}$ is the UIP from the conflicting
  level. $\stablealluip$ first tries to find the UIP for level 6 by
  resolving $C_1$ with $C_k$, $C_j$ and then $C_i$ producing the
  clause $C^* = (\neg{m_{10}}$, $\neg{h_6}$, $\neg{g_5}$, $\neg{f_5}$,
  $\neg{e_5}$, $\neg{d_2}$, $\neg{c_2})$ where $\neg{h_6}$ is the UIP
  for level 6.

@@

\subsubsection{$\allUipAct$: Clauses with Active Variables.}
\label{sec:active} $\stablealluip$ learning might introduce literals
with low variable activity into $C_i$. Low activity variables are
variables that have had low recent participation in clause
learning. Hence, clauses with variables of low activity might not be
as currently useful to the solver. Our variant $\allUipAct$
(line~\ref{ln:activity} in Algorithm~\ref{alg:i-uip}) computes the
average variable activity of the newly produced \alluip clause $C_i$
and the original \oneuip clause $C_1$. The new clause $C_i$ will be
returned only if it is both smaller and has higher average variable
activity than the original \oneuip clause. There are, of course,
generalizations of this approach where one has a weighted trade-off
between these factors that allows preferring the new clause when it
has large gains in one metric even though it has small losses in the
other. We did not, however, experiment with such generalizations.
